### 2020-04-06 submit2/makeFigs/multiFluidBubble_500_1.tex

%% LyX 2.2.4 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[12pt,british,times]{article}
\usepackage{mathptmx}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage[pdftex,a4paper]{geometry}
\geometry{verbose,tmargin=2cm,bmargin=2cm,lmargin=2cm,rmargin=2cm}
\pagestyle{empty}
\setcounter{secnumdepth}{5}
\setcounter{tocdepth}{5}
\setlength{\parskip}{\smallskipamount}
\setlength{\parindent}{0pt}
\usepackage{calc}
\usepackage{amstext}
\usepackage{amssymb}
\usepackage[pdftex]{graphicx}
\usepackage[authoryear]{natbib}

\makeatletter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
%% Because html converters don't know tabularnewline
\providecommand{\tabularnewline}{\\}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
\usepackage { fancybox}
\usepackage[pdftex]{graphicx,color}
\usepackage{rotating}
\newcommand{\nicefrac}[2]{\ensuremath ^{#1}\!\!/\!_{#2}}
\usepackage { fancybox}
\usepackage[export]{adjustbox}
\usepackage{amsmath}

\makeatother

\usepackage{babel}
\begin{document}
\begin{tabular}{cccc}
\multicolumn{4}{c}{%
\fbox{\begin{minipage}[t]{0.7\columnwidth}%
\begin{center}
$S_{ij}=0$, \hspace{6pt} $\gamma=0$, \hspace{6pt} $C_{D}/r_{c}=0$,
\hspace{6pt} $b_{ij}^{T}=b_{i}$
\par\end{center}%
\end{minipage}}}\tabularnewline
\includegraphics[scale=0.6]{/home/hilary/OpenFOAM/hilary-7/run/hilary/warmBubble/Boussinesq/multiFluid/singleColumn_noTransfer/500/sigmaCompare} & \includegraphics[scale=0.6]{/home/hilary/OpenFOAM/hilary-7/run/hilary/warmBubble/Boussinesq/multiFluid/singleColumn_noTransfer/500/bCompare} & \includegraphics[scale=0.6]{/home/hilary/OpenFOAM/hilary-7/run/hilary/warmBubble/Boussinesq/multiFluid/singleColumn_noTransfer/500/wCompare} & \includegraphics[scale=0.6]{/home/hilary/OpenFOAM/hilary-7/run/hilary/warmBubble/Boussinesq/multiFluid/singleColumn_noTransfer/500/Pcompare}\tabularnewline
\multicolumn{4}{c}{%
\fbox{\begin{minipage}[t]{0.7\columnwidth}%
\begin{center}
$S_{ij}=-\nabla\cdot\mathbf{u}_{i}$, \hspace{6pt} $\gamma=0$, \hspace{6pt}
$C_{D}/r_{c}=0$, \hspace{6pt} $b_{ij}^{T}=b_{i}$
\par\end{center}%
\end{minipage}}}\tabularnewline
\includegraphics[scale=0.6]{/home/hilary/OpenFOAM/hilary-7/run/hilary/warmBubble/Boussinesq/multiFluid/singleColumn_divTransfer/500/sigmaCompare} & \includegraphics[scale=0.6]{/home/hilary/OpenFOAM/hilary-7/run/hilary/warmBubble/Boussinesq/multiFluid/singleColumn_divTransfer/500/bCompare} & \includegraphics[scale=0.6]{/home/hilary/OpenFOAM/hilary-7/run/hilary/warmBubble/Boussinesq/multiFluid/singleColumn_divTransfer/500/wCompare} & \includegraphics[scale=0.6]{/home/hilary/OpenFOAM/hilary-7/run/hilary/warmBubble/Boussinesq/multiFluid/singleColumn_divTransfer/500/Pcompare}\tabularnewline
\multicolumn{4}{c}{%
\fbox{\begin{minipage}[t]{0.7\columnwidth}%
\begin{center}
$S_{ij}=-\nabla\cdot\mathbf{u}_{i}$, \hspace{6pt} $\gamma=2.3\!\!\times\!\!10^{4}\text{m}^{2}\text{s}^{-1}$,
\hspace{6pt} $C_{D}/r_{c}=0$, \hspace{6pt} $b_{ij}^{T}=b_{i}$
\par\end{center}%
\end{minipage}}}\tabularnewline
\includegraphics[scale=0.6]{/home/hilary/OpenFOAM/hilary-7/run/hilary/warmBubble/Boussinesq/multiFluid/singleColumn_Pi_2p3e4_divTransfer/500/sigmaCompare} & \includegraphics[scale=0.6]{/home/hilary/OpenFOAM/hilary-7/run/hilary/warmBubble/Boussinesq/multiFluid/singleColumn_Pi_2p3e4_divTransfer/500/bCompare} & \includegraphics[scale=0.6]{/home/hilary/OpenFOAM/hilary-7/run/hilary/warmBubble/Boussinesq/multiFluid/singleColumn_Pi_2p3e4_divTransfer/500/wCompare} & \includegraphics[scale=0.6]{/home/hilary/OpenFOAM/hilary-7/run/hilary/warmBubble/Boussinesq/multiFluid/singleColumn_Pi_2p3e4_divTransfer/500/Pcompare}\tabularnewline
\multicolumn{4}{c}{%
\fbox{\begin{minipage}[t]{0.7\columnwidth}%
\begin{center}
$S_{ij}=-\nabla\cdot\mathbf{u}_{i}$, \hspace{6pt} $\gamma=4\!\!\times\!\!10^{4}\text{m}^{2}\text{s}^{-1}$,
\hspace{6pt} $C_{D}/r_{c}=0$, \hspace{6pt} $b_{ij}^{T}=b_{i}$
\par\end{center}%
\end{minipage}}}\tabularnewline
\includegraphics[scale=0.6]{/home/hilary/OpenFOAM/hilary-7/run/hilary/warmBubble/Boussinesq/multiFluid/singleColumn_Pi_4e4_divTransfer/500/sigmaCompare} & \includegraphics[scale=0.6]{/home/hilary/OpenFOAM/hilary-7/run/hilary/warmBubble/Boussinesq/multiFluid/singleColumn_Pi_4e4_divTransfer/500/bCompare} & \includegraphics[scale=0.6]{/home/hilary/OpenFOAM/hilary-7/run/hilary/warmBubble/Boussinesq/multiFluid/singleColumn_Pi_4e4_divTransfer/500/wCompare} & \includegraphics[scale=0.6]{/home/hilary/OpenFOAM/hilary-7/run/hilary/warmBubble/Boussinesq/multiFluid/singleColumn_Pi_4e4_divTransfer/500/Pcompare}\tabularnewline
\multicolumn{4}{c}{%
\fbox{\begin{minipage}[t]{0.7\columnwidth}%
\begin{center}
$S_{ij}=-\nabla\cdot\mathbf{u}_{i}$, \hspace{6pt} $\gamma=6\!\!\times\!\!10^{4}\text{m}^{2}\text{s}^{-1}$,
\hspace{6pt} $C_{D}/r_{c}=0$, \hspace{6pt} $b_{ij}^{T}=b_{i}$
\par\end{center}%
\end{minipage}}}\tabularnewline
\includegraphics[scale=0.6]{/home/hilary/OpenFOAM/hilary-7/run/hilary/warmBubble/Boussinesq/multiFluid/singleColumn_Pi_6e4_divTransfer/500/sigmaCompare} & \includegraphics[scale=0.6]{/home/hilary/OpenFOAM/hilary-7/run/hilary/warmBubble/Boussinesq/multiFluid/singleColumn_Pi_6e4_divTransfer/500/bCompare} & \includegraphics[scale=0.6]{/home/hilary/OpenFOAM/hilary-7/run/hilary/warmBubble/Boussinesq/multiFluid/singleColumn_Pi_6e4_divTransfer/500/wCompare} & \includegraphics[scale=0.6]{/home/hilary/OpenFOAM/hilary-7/run/hilary/warmBubble/Boussinesq/multiFluid/singleColumn_Pi_6e4_divTransfer/500/Pcompare}\tabularnewline
$\sigma_{1}$ & $b$, $b_{0}$ and $b_{1}$ & $w_{0}$ and $w_{1}$ & $P_{0}$ and $P_{1}$\tabularnewline
\end{tabular}
\end{document}